\chapter{《Attention Is All You Need》中文版}

\section{摘要}

目前主流的序列转换模型建立在复杂的循环神经网络或卷积神经网络之上，通常包含一个编码器和一个解码器。性能最优的模型还通过注意力机制将编码器与解码器连接起来。本文提出一种全新的简单网络架构——Transformer，它完全基于注意力机制，彻底摒弃了循环结构与卷积操作。在两项机器翻译任务上的实验表明，该模型在翻译质量上更优，同时具备更强的并行性，所需训练时间也大幅缩短。在 WMT 2014 英德翻译任务上，我们的模型取得了 28.4 的 BLEU 分数，比现有最优结果（包括集成模型）提升了 2 个 BLEU 以上。在 WMT 2014 英法翻译任务上，我们的模型仅用 8 块 GPU 训练 3.5 天，便以单模型创下 41.8 的 BLEU 新纪录，训练成本仅为文献中最优模型的一小部分。此外，我们将 Transformer 成功应用于英语成分句法分析任务（无论训练数据规模大小），证明其具备良好的泛化能力。

\section{引言}

循环神经网络，尤其是长短期记忆网络（LSTM）\cite{hochreiter1997long}和门控循环神经网络\cite{cho2014learning}，已在序列建模与转换问题（如语言建模和机器翻译）中牢固确立了其最先进方法的地位\cite{sutskever2014sequence,cho2014learning,bahdanau2014neural}。此后，大量研究持续推动着循环语言模型与编码器-解码器架构的边界\cite{wu2016google,luong2015effective,jozefowicz2016exploring}。

循环模型通常沿输入和输出序列的符号位置逐步推进计算。将位置与计算时间步对齐后，模型以上一时刻的隐藏状态 $h_{t-1}$ 和当前位置 $t$ 的输入为函数，生成一系列隐藏状态 $h_t$。这种本质上的顺序性使得训练样本内部无法并行化，在序列较长时尤为制约性能，因为内存限制会压缩跨样本的批处理规模。近期工作通过分解技巧\cite{kuchaiev2017factorization}和条件计算\cite{shazeer2017outrageously}在计算效率上取得了显著提升，后者同时也改善了模型性能。然而，顺序计算这一根本约束依然存在。

注意力机制已成为各类序列建模与转换模型中不可或缺的组成部分，它允许建模输入或输出序列中任意距离的依赖关系\cite{bahdanau2014neural,structuredattentionnetworks}。然而，除少数例外\cite{decomposableattentionmodel}，注意力机制几乎都与循环网络配合使用。

本文提出 Transformer——一种完全摒弃循环结构、仅依赖注意力机制来捕捉输入与输出之间全局依赖关系的模型架构。Transformer 具备更强的并行化能力，仅需在 8 块 P100 GPU 上训练短短十二小时，便可在翻译质量上达到新的最先进水平。

\section{背景}

减少顺序计算这一目标，同样是 Extended Neural GPU\cite{extendedneuralgpu}、ByteNet\cite{kalchbrenner2016neural} 和 ConvS2S\cite{gehring2017convolutional} 的出发点。这些模型均以卷积神经网络为基本构件，能够并行计算所有输入和输出位置的隐藏表示。但在这些模型中，关联两个任意输入或输出位置信号所需的操作次数，随位置间距的增大而增长——ConvS2S 呈线性增长，ByteNet 呈对数增长。这使得学习远距离位置之间的依赖关系更加困难\cite{hochreiter2001gradient}。在 Transformer 中，这一代价被降低为常数次操作，尽管代价是因对注意力权重位置取平均而导致有效分辨率下降——我们通过第 3.2 节介绍的多头注意力机制来抵消这一影响。

自注意力（有时称为内部注意力）是一种将单个序列中不同位置相互关联、从而计算序列表示的注意力机制。自注意力已在阅读理解、抽象摘要、文本蕴含及与任务无关的句子表示学习等多种任务中取得成功\cite{cheng2016long,decomposableattentionmodel,paulus2017deep,lin2017structured}。

端到端记忆网络基于循环注意力机制（而非序列对齐的循环结构），已被证明在简单语言问答和语言建模任务上表现良好\cite{sukhbaatar2015}。

然而，据我们所知，Transformer 是首个完全依赖自注意力来计算输入和输出表示、不使用任何序列对齐 RNN 或卷积的转换模型。在接下来的章节中，我们将描述 Transformer，阐述自注意力的动机，并讨论其相较于\cite{sutskever2014sequence,bahdanau2014neural}等模型的优势。

\section{模型架构}

大多数具有竞争力的神经序列转换模型都采用编码器-解码器结构\cite{cho2014learning,bahdanau2014neural,sutskever2014sequence}。编码器将输入符号表示序列 $(x_1, \ldots, x_n)$ 映射为连续表示序列 $\mathbf{z} = (z_1, \ldots, z_n)$。给定 $\mathbf{z}$ 后，解码器逐步生成输出符号序列 $(y_1, \ldots, y_m)$，每次生成一个元素。在每一步中，模型是自回归的\cite{graves2013generating}，即在生成下一个符号时，将已生成的符号作为额外输入。

Transformer 遵循上述整体架构，对编码器和解码器均采用堆叠的自注意力层与逐点全连接层，分别对应图 1 的左半部分和右半部分。

\subsection{编码器与解码器堆栈}

\textbf{编码器：}编码器由 $N = 6$ 个相同层堆叠而成。每层包含两个子层：第一个是多头自注意力机制，第二个是简单的逐位置全连接前馈网络。在每个子层周围采用残差连接\cite{he2016deep}，随后进行层归一化\cite{ba2016layer}。即每个子层的输出为 $\mathrm{LayerNorm}(x + \mathrm{Sublayer}(x))$，其中 $\mathrm{Sublayer}(x)$ 为该子层自身实现的函数。为便于使用残差连接，模型中所有子层及嵌入层的输出维度均为 $d_{\text{model}} = 512$。

\textbf{解码器：}解码器同样由 $N = 6$ 个相同层堆叠而成。在编码器每层的两个子层基础上，解码器额外插入第三个子层，对编码器堆栈的输出执行多头注意力。与编码器类似，每个子层周围同样采用残差连接并进行层归一化。此外，对解码器堆栈中的自注意力子层进行了修改，通过掩码（Masking）防止当前位置关注到后续位置。这种掩码结合输出嵌入偏移一个位置的设计，确保位置 $i$ 的预测只能依赖位置 $i$ 之前的已知输出。

\subsection{注意力}

注意力函数可描述为：将一个查询（query）和一组键值对（key-value pairs）映射为输出，其中查询、键、值和输出均为向量。输出是值的加权求和，每个值的权重由查询与对应键的相容性函数计算得到。

\subsubsection{缩放点积注意力}

我们将本文的注意力称为"缩放点积注意力"（图 2）。输入包含维度为 $d_k$ 的查询和键，以及维度为 $d_v$ 的值。将查询与所有键做点积，除以 $\sqrt{d_k}$ 后，应用 softmax 函数得到各值的权重。

在实践中，我们将一组查询打包成矩阵 $Q$，同时将键和值分别打包成矩阵 $K$ 和 $V$，并行计算注意力函数，输出矩阵为：
\[
  \mathrm{Attention}(Q, K, V) = \mathrm{softmax}\!\left(\frac{QK^{\top}}{\sqrt{d_k}}\right)V
\]

最常用的两种注意力函数是加性注意力\cite{bahdanau2014neural}和点积（乘法）注意力。点积注意力与本文算法完全相同，区别仅在于缩放因子 $\frac{1}{\sqrt{d_k}}$。加性注意力使用单隐层前馈网络计算相容性函数。二者理论复杂度相近，但点积注意力在实践中速度更快、空间效率更高，因为可以借助高度优化的矩阵乘法代码实现。

当 $d_k$ 较小时，两种机制性能相当；当 $d_k$ 较大时，不加缩放的点积注意力性能逊于加性注意力\cite{DBLP:journals/corr/BritzGLL17}。我们推测，当 $d_k$ 较大时，点积结果幅值增大，将 softmax 函数推入梯度极小的区域。为抵消这一效应，我们将点积结果除以 $\frac{1}{\sqrt{d_k}}$ 进行缩放。

\subsubsection{多头注意力}

我们发现，与其用 $d_{\text{model}}$ 维的键、值和查询执行单次注意力函数，不如将查询、键和值分别经过 $h$ 次不同的可学习线性投影，投影到 $d_k$、$d_k$ 和 $d_v$ 维度，效果更佳。在每组投影后的查询、键和值上并行执行注意力函数，得到 $d_v$ 维输出值，将其拼接后再次投影，得到最终输出，如图 2 所示。

多头注意力使模型能够在不同位置上同时关注来自不同表示子空间的信息，而单头注意力的平均化操作会抑制这一能力。
\[
  \mathrm{MultiHead}(Q, K, V) = \mathrm{Concat}(\mathrm{head}_1, \ldots, \mathrm{head}_h)\,W^O
\]
其中 $\mathrm{head}_i = \mathrm{Attention}(QW_i^Q,\, KW_i^K,\, VW_i^V)$

投影矩阵参数为
$W_i^Q \in \mathbb{R}^{d_{\text{model}} \times d_k}$，
$W_i^K \in \mathbb{R}^{d_{\text{model}} \times d_k}$，
$W_i^V \in \mathbb{R}^{d_{\text{model}} \times d_v}$，
$W^O \in \mathbb{R}^{hd_v \times d_{\text{model}}}$。

本文采用 $h = 8$ 个并行注意力头，每个头的维度为 $d_k = d_v = d_{\text{model}}/h = 64$。由于每个头的维度缩减，总计算量与全维度单头注意力相近。

\subsubsection{注意力在模型中的应用}

Transformer 以三种不同方式使用多头注意力：

\begin{itemize}
  \item 在"编码器-解码器注意力"层中，查询来自上一解码器层，键和值来自编码器的输出。这使解码器的每个位置都能关注输入序列的所有位置，与 seq2seq 模型中的典型编码器-解码器注意力机制\cite{wu2016google,bahdanau2014neural,decomposableattentionmodel}类似。
  \item 编码器包含自注意力层。在自注意力层中，键、值和查询均来自同一处——上一编码器层的输出。编码器中每个位置都可以关注编码器上一层的所有位置。
  \item 同样，解码器中的自注意力层允许每个位置关注解码器中该位置及之前的所有位置。为保持自回归性质，需要阻止解码器中信息向左流动。我们在缩放点积注意力内部实现这一点：将 softmax 输入中对应非法连接的所有值掩码为 $-\infty$。
\end{itemize}

\subsection{逐位置前馈网络}

除注意力子层外，编码器和解码器的每一层还包含一个全连接前馈网络，对每个位置分别且相同地应用。该网络由两个线性变换组成，中间加入 ReLU 激活函数：
\[
  \mathrm{FFN}(x) = \max(0,\, xW_1 + b_1)\,W_2 + b_2
\]
不同位置的线性变换相同，但不同层之间使用不同参数。也可将其理解为两个核大小为 1 的卷积。输入和输出维度为 $d_{\text{model}} = 512$，内层维度为 $d_{ff} = 2048$。

\subsection{嵌入与 Softmax}

与其他序列转换模型类似，我们使用可学习的嵌入将输入词元和输出词元转换为 $d_{\text{model}}$ 维向量。同时使用常规的可学习线性变换和 softmax 函数，将解码器输出转换为预测下一词元的概率。本模型在两个嵌入层与 pre-softmax 线性变换之间共享权重矩阵\cite{press2016using}，在嵌入层中将权重乘以 $\sqrt{d_{\text{model}}}$。

\subsection{位置编码}

由于本模型不含循环结构也不含卷积，为使模型能够利用序列的顺序信息，必须向模型注入词元在序列中的相对或绝对位置信息。为此，我们在编码器和解码器堆栈底部的输入嵌入上叠加"位置编码"。位置编码与嵌入的维度相同，均为 $d_{\text{model}}$，因此二者可以直接相加。位置编码有多种选择，包括可学习的和固定的\cite{gehring2017convolutional}。

本文使用不同频率的正弦和余弦函数：
\[
  PE_{(pos,\,2i)}   = \sin\!\left(\frac{pos}{10000^{2i/d_{\text{model}}}}\right)
  \qquad
  PE_{(pos,\,2i+1)} = \cos\!\left(\frac{pos}{10000^{2i/d_{\text{model}}}}\right)
\]
其中 $pos$ 为位置，$i$ 为维度索引。每个维度的位置编码对应一条正弦曲线，波长从 $2\pi$ 到 $10000 \cdot 2\pi$ 构成等比数列。选择该函数是因为我们假设它能使模型轻松学会通过相对位置进行注意力计算——对于任意固定偏移 $k$，$PE_{pos+k}$ 可表示为 $PE_{pos}$ 的线性函数。

我们还实验了使用可学习的位置嵌入\cite{gehring2017convolutional}，发现两种方式结果几乎相同（见表 3 第 (E) 行）。最终选择正弦版本，因为它可能允许模型外推至训练时未见过的更长序列。

\section{为什么使用自注意力}

本节将自注意力层与循环层、卷积层在多个方面进行比较。这些层通常用于将一段变长符号表示序列 $(x_1, \ldots, x_n)$ 映射为另一段等长序列 $(z_1, \ldots, z_n)$，其中 $x_i, z_i \in \mathbb{R}^d$，例如典型序列转换编码器或解码器中的隐藏层。我们从三个方面论证使用自注意力的动机。

其一是每层的总计算复杂度。其二是可并行化的计算量，以所需最少顺序操作数衡量。

其三是网络中长距离依赖之间的路径长度。学习长距离依赖是许多序列转换任务的核心挑战。影响这一能力的关键因素之一，是前向与反向信号在网络中需要经过的路径长度。输入与输出序列中任意位置组合之间的路径越短，越容易学习长距离依赖\cite{hochreiter2001gradient}。因此，我们也比较了由不同类型层构成的网络中，任意两个输入输出位置之间的最大路径长度。

如表 1 所示，自注意力层以常数次顺序操作连接所有位置，而循环层需要 $O(n)$ 次顺序操作。在计算复杂度方面，当序列长度 $n$ 小于表示维度 $d$ 时，自注意力层比循环层更快——这在机器翻译中最先进模型所使用的句子表示（如词片\cite{wu2016google}和字节对\cite{sennrich2015neural}表示）中最为常见。

为提升处理超长序列任务时的计算性能，自注意力可限制为只关注以各输出位置为中心、大小为 $r$ 的邻域。这将使最大路径长度增至 $O(n/r)$，我们计划在未来工作中进一步研究这一方法。

核宽为 $k < n$ 的单层卷积层无法连接所有输入输出位置对。若要做到这一点，连续核情况下需要堆叠 $O(n/k)$ 层卷积，膨胀卷积\cite{yu2015multi}情况下需要 $O(\log_k(n))$ 层，这都会增加网络中任意两个位置之间最长路径的长度。卷积层的计算代价通常比循环层高出 $k$ 倍。然而，可分离卷积\cite{chollet2016xception}可将复杂度大幅降低至 $O(k \cdot n \cdot d + n \cdot d^2)$。即便取 $k = n$，可分离卷积的复杂度也与自注意力层加逐点前馈层的组合相当——正是本文模型所采用的方案。

自注意力还有一个附带好处：可产生更具可解释性的模型。我们对模型的注意力分布进行了检视，并在附录中列举和讨论了若干示例。不仅各个注意力头明显学会了执行不同的任务，许多头还表现出与句子句法和语义结构相关的行为。

\section{训练}

本节介绍模型的训练方案。

\subsection{训练数据与批处理}

我们在标准的 WMT 2014 英德数据集上进行训练，该数据集约包含 450 万个句子对。句子采用字节对编码\cite{sennrich2015neural}，源语言与目标语言共享约 37000 个词元的词表。对于英法翻译，我们使用规模更大的 WMT 2014 英法数据集，共 3600 万个句子，词元切分为 32000 个词片词表\cite{wu2016google}。句子对按近似序列长度打包成批次，每个训练批次约包含 25000 个源词元和 25000 个目标词元。

\subsection{硬件与训练计划}

我们在一台配备 8 块 NVIDIA P100 GPU 的机器上训练模型。使用本文所描述超参数的基础模型，每个训练步耗时约 0.4 秒，共训练 100000 步，历时约 12 小时。大模型每步耗时 1.0 秒，共训练 300000 步，历时 3.5 天。

\subsection{优化器}

我们使用 Adam 优化器\cite{kingma2014adam}，参数设置为 $\beta_1 = 0.9$，$\beta_2 = 0.98$，$\epsilon = 10^{-9}$。训练过程中按以下公式动态调整学习率：
\[
  lrate = d_{\text{model}}^{-0.5} \cdot
  \min\!\left(step\_num^{-0.5},\;
  step\_num \cdot warmup\_steps^{-1.5}\right)
\]
即在前 $warmup\_steps$ 步线性增大学习率，之后按步数的平方根倒数衰减。本文设 $warmup\_steps = 4000$。

\subsection{正则化}

训练中采用三种正则化方式：

\textbf{残差 Dropout。}在每个子层的输出加到子层输入并归一化之前，对其施加 Dropout\cite{srivastava2014dropout}。此外，在编码器和解码器堆栈中，对嵌入与位置编码之和也施加 Dropout。基础模型的丢弃率为 $P_{drop} = 0.1$。

\textbf{标签平滑。}训练时采用值为 $\epsilon_{ls} = 0.1$ 的标签平滑\cite{szegedy2015rethinking}。这会使模型变得更加"不确定"，从而损害困惑度，但能提升准确率和 BLEU 分数。

\section{实验结果}

\subsection{机器翻译}

在 WMT 2014 英德翻译任务上，大型 Transformer 模型（表 2 中的 Transformer (big)）以超过 2.0 BLEU 的优势超越了此前所有已报告的最优模型（包括集成模型），以 28.4 的 BLEU 分数创下新的最先进水平。该模型的配置见表 3 最后一行，在 8 块 P100 GPU 上训练耗时 3.5 天。即便是我们的基础模型，也以远低于任何竞争模型的训练成本超越了此前所有已发表的模型和集成模型。

在 WMT 2014 英法翻译任务上，我们的大模型取得了 41.0 的 BLEU 分数，超越了此前所有已发表的单模型，训练成本不足此前最优模型的四分之一。英法任务的大型 Transformer 模型使用的 Dropout 率为 $P_{drop} = 0.1$，而非 0.3。

对于基础模型，我们对最后 5 个检查点（每 10 分钟保存一次）取平均，得到单一模型。对于大模型，则对最后 20 个检查点取平均。推理时使用束搜索，束大小为 4，长度惩罚系数 $\alpha = 0.6$\cite{wu2016google}，这些超参数均经过开发集实验选定。推理时最大输出长度设为输入长度加 50，但在可能时提前终止\cite{wu2016google}。

表 2 汇总了我们的实验结果，并将翻译质量和训练成本与文献中其他模型架构进行了比较。训练模型所用的浮点运算次数，由训练时间、所用 GPU 数量以及每块 GPU 的单精度浮点运算峰值估算得出。

\subsection{模型变体}

为评估 Transformer 各组件的重要性，我们以不同方式对基础模型进行变体实验，在开发集 newstest2013 上衡量英德翻译性能的变化，结果见表 3。推理使用上节所述的束搜索，但不进行检查点平均。

在表 3 第 (A) 组，我们按第 3.2.2 节的描述，在保持计算量不变的前提下，改变注意力头数及注意力键值维度。结果显示，单头注意力比最优设置差 0.9 BLEU，但头数过多时质量同样下降。

在第 (B) 组，我们观察到减小注意力键维度 $d_k$ 会损害模型质量，这表明判断相容性并不容易，比点积更复杂的相容性函数或许会更有益。在第 (C) 和 (D) 组，进一步验证了更大的模型效果更好，Dropout 对于防止过拟合非常有效。在第 (E) 行，用可学习的位置嵌入\cite{gehring2017convolutional}替换正弦位置编码后，结果与基础模型几乎相同。

\subsection{英语成分句法分析}

为评估 Transformer 能否泛化到其他任务，我们在英语成分句法分析任务上进行了实验。该任务有其特殊挑战：输出受到强结构约束，且通常远长于输入；此外，在小数据场景下，RNN seq2seq 模型难以达到最先进水平\cite{vinyals2015grammar}。

我们在 Penn Treebank\cite{marcus1993building} 的华尔街日报（WSJ）部分（约 4 万训练句）上训练了一个 4 层、$d_{\text{model}} = 1024$ 的 Transformer。同时还在半监督设置下进行了训练，利用了约 1700 万句的高置信度语料和 BerkeleyParser 语料\cite{vinyals2015grammar,zhu2013fast}。纯 WSJ 设置使用 16K 词元词表，半监督设置使用 32K 词元词表。

我们仅在第 22 节开发集上进行了少量实验以选择 Dropout 率（注意力和残差，见第 5.4 节）、学习率和束大小，其余所有参数均沿用英德基础翻译模型的设置。推理时最大输出长度设为输入长度加 300，束大小为 21，$\alpha = 0.3$（纯 WSJ 和半监督设置相同）。

表 4 的结果表明，尽管缺乏针对特定任务的调优，我们的模型表现出乎意料地好，优于此前所有已报告模型，仅次于循环神经网络文法模型（RNNG）\cite{dyer2016recurrent}。与 RNN seq2seq 模型\cite{vinyals2015grammar}不同，即使仅在 4 万句的 WSJ 训练集上训练，Transformer 也超越了 BerkeleyParser\cite{petrov2006learning}。

\section{结论}

本文提出了 Transformer——首个完全基于注意力机制的序列转换模型，以多头自注意力取代了编码器-解码器架构中最常用的循环层。

在翻译任务上，Transformer 的训练速度显著快于基于循环或卷积层的架构。在 WMT 2014 英德和英法翻译任务上均取得了新的最先进水平，前者甚至超越了此前所有已报告的集成模型。

我们对基于注意力模型的未来充满期待，并计划将其应用于其他任务。我们还计划将 Transformer 扩展到文本以外的输入输出模态，并研究局部受限注意力机制，以高效处理图像、音频和视频等大规模输入输出。减少生成过程的顺序性也是我们的研究目标之一。

本文训练和评估模型所用的代码已开源，地址为 \url{https://github.com/tensorflow/tensor2tensor}。

\chapter{Attention Is All You Need}

\section{Abstract}

The dominant sequence transduction models are based on complex recurrent or
convolutional neural networks that include an encoder and a decoder. The best
performing models also connect the encoder and decoder through an attention
mechanism. We propose a new simple network architecture, the Transformer,
based solely on attention mechanisms, dispensing with recurrence and
convolutions entirely. Experiments on two machine translation tasks show these
models to be superior in quality while being more parallelizable and requiring
significantly less time to train. Our model achieves 28.4 BLEU on the WMT
2014 English-to-German translation task, improving over the existing best
results, including ensembles, by over 2 BLEU. On the WMT 2014
English-to-French translation task, our model establishes a new single-model
state-of-the-art BLEU score of 41.8 after training for 3.5 days on eight
GPUs, a small fraction of the training costs of the best models from the
literature. We show that the Transformer generalizes well to other tasks by
applying it successfully to English constituency parsing both with large and
limited training data.

\section{Introduction}

Recurrent neural networks, long short-term memory \cite{hochreiter1997long}
and gated recurrent \cite{cho2014learning} neural networks in particular, have
been firmly established as state of the art approaches in sequence modeling
and transduction problems such as language modeling and machine translation
\cite{sutskever2014sequence,cho2014learning,bahdanau2014neural}. Numerous
efforts have since continued to push the boundaries of recurrent language
models and encoder-decoder architectures
\cite{wu2016google,luong2015effective,jozefowicz2016exploring}.

Recurrent models typically factor computation along the symbol positions of
the input and output sequences. Aligning the positions to steps in computation
time, they generate a sequence of hidden states $h_t$, as a function of the
previous hidden state $h_{t-1}$ and the input for position $t$. This inherently
sequential nature precludes parallelization within training examples, which
becomes critical at longer sequence lengths, as memory constraints limit
batching across examples. Recent work has achieved significant improvements in
computational efficiency through factorization tricks
\cite{kuchaiev2017factorization} and conditional computation
\cite{shazeer2017outrageously}, while also improving model performance in case
of the latter. The fundamental constraint of sequential computation, however,
remains.

Attention mechanisms have become an integral part of compelling sequence
modeling and transduction models in various tasks, allowing modeling of
dependencies without regard to their distance in the input or output sequences
\cite{bahdanau2014neural,structuredattentionnetworks}. In all but a few cases
\cite{decomposableattentionmodel}, however, such attention mechanisms are used
in conjunction with a recurrent network.

In this work we propose the Transformer, a model architecture eschewing
recurrence and instead relying entirely on an attention mechanism to draw
global dependencies between input and output. The Transformer allows for
significantly more parallelization and can reach a new state of the art in
translation quality after being trained for as little as twelve hours on eight
P100 GPUs.

\section{Background}

The goal of reducing sequential computation also forms the foundation of the
Extended Neural GPU \cite{extendedneuralgpu}, ByteNet
\cite{kalchbrenner2016neural} and ConvS2S \cite{gehring2017convolutional}, all
of which use convolutional neural networks as basic building block, computing
hidden representations in parallel for all input and output positions. In
these models, the number of operations required to relate signals from two
arbitrary input or output positions grows in the distance between positions,
linearly for ConvS2S and logarithmically for ByteNet. This makes it more
difficult to learn dependencies between distant positions
\cite{hochreiter2001gradient}. In the Transformer this is reduced to a
constant number of operations, albeit at the cost of reduced effective
resolution due to averaging attention-weighted positions, an effect we
counteract with Multi-Head Attention as described in Section~3.2.

Self-attention, sometimes called intra-attention is an attention mechanism
relating different positions of a single sequence in order to compute a
representation of the sequence. Self-attention has been used successfully in a
variety of tasks including reading comprehension, abstractive summarization,
textual entailment and learning task-independent sentence representations
\cite{cheng2016long,decomposableattentionmodel,paulus2017deep,lin2017structured}.

End-to-end memory networks are based on a recurrent attention mechanism
instead of sequence-aligned recurrence and have been shown to perform well on
simple-language question answering and language modeling tasks
\cite{sukhbaatar2015}.

To the best of our knowledge, however, the Transformer is the first transduction
model relying entirely on self-attention to compute representations of its input
and output without using sequence-aligned RNNs or convolution. In the following
sections, we will describe the Transformer, motivate self-attention and discuss
its advantages over models such as
\cite{sutskever2014sequence,bahdanau2014neural} and
\cite{recurrentleastsquares,lecun2015deep}.

\section{Model Architecture}

Most competitive neural sequence transduction models have an encoder-decoder
structure \cite{cho2014learning,bahdanau2014neural,sutskever2014sequence}. Here,
the encoder maps an input sequence of symbol representations
$(x_1, \ldots, x_n)$ to a sequence of continuous representations
$\mathbf{z} = (z_1, \ldots, z_n)$. Given $\mathbf{z}$, the decoder then
generates an output sequence $(y_1, \ldots, y_m)$ of symbols one element at a
time. At each step the model is auto-regressive \cite{graves2013generating},
consuming the previously generated symbols as additional input when generating
the next.

The Transformer follows this overall architecture using stacked self-attention
and point-wise, fully connected layers for both the encoder and decoder, shown
in the left and right halves of Figure~1, respectively.

\subsection{Encoder and Decoder Stacks}

\textbf{Encoder:} The encoder is composed of a stack of $N = 6$ identical
layers. Each layer has two sub-layers. The first is a multi-head self-attention
mechanism, and the second is a simple, position-wise fully connected
feed-forward network. We employ a residual connection
\cite{he2016deep} around each of the two sub-layers, followed by layer
normalization \cite{ba2016layer}. That is, the output of each sub-layer is
$\mathrm{LayerNorm}(x + \mathrm{Sublayer}(x))$, where $\mathrm{Sublayer}(x)$
is the function implemented by the sub-layer itself. To facilitate these
residual connections, all sub-layers in the model, as well as the embedding
layers, produce outputs of dimension $d_{\text{model}} = 512$.

\textbf{Decoder:} The decoder is also composed of a stack of $N = 6$ identical
layers. In addition to the two sub-layers in each encoder layer, the decoder
inserts a third sub-layer, which performs multi-head attention over the output
of the encoder stack. Similar to the encoder, we employ residual connections
around each of the sub-layers, followed by layer normalization. We also modify
the self-attention sub-layer in the decoder stack to prevent positions from
attending to subsequent positions. This masking, combined with fact that the
output embeddings are offset by one position, ensures that the predictions for
position $i$ can depend only on the known outputs at positions less than $i$.

\subsection{Attention}

An attention function can be described as mapping a query and a set of
key-value pairs to an output, where the query, keys, values, and output are
all vectors. The output is computed as a weighted sum of the values, where the
weight assigned to each value is computed by a compatibility function of the
query with the corresponding key.

\subsubsection{Scaled Dot-Product Attention}

We call our particular attention ``Scaled Dot-Product Attention'' (Figure~2).
The input consists of queries and keys of dimension $d_k$, and values of
dimension $d_v$. We compute the dot products of the query with all keys, divide
each by $\sqrt{d_k}$, and apply a softmax function to obtain the weights on
the values.

In practice, we compute the attention function on a set of queries
simultaneously, packed together into a matrix $Q$. The keys and values are also
packed together into matrices $K$ and $V$. We compute the matrix of outputs as:
\[
  \mathrm{Attention}(Q, K, V) = \mathrm{softmax}\!\left(\frac{QK^{\top}}{\sqrt{d_k}}\right)V
\]

The two most commonly used attention functions are additive attention
\cite{bahdanau2014neural}, and dot-product (multiplicative) attention.
Dot-product attention is identical to our algorithm, except for the scaling
factor of $\frac{1}{\sqrt{d_k}}$. Additive attention computes the compatibility
function using a feed-forward network with a single hidden layer. While the two
are similar in theoretical complexity, dot-product attention is much faster and
more space-efficient in practice, since it can be implemented using highly
optimized matrix multiplication code.

While for small values of $d_k$ the two mechanisms perform similarly, additive
attention outperforms dot product attention without scaling for larger values of
$d_k$ \cite{DBLP:journals/corr/BritzGLL17}. We suspect that for large values
of $d_k$, the dot products grow large in magnitude, pushing the softmax function
into regions where it has extremely small gradients. To counteract this effect,
we scale the dot products by $\frac{1}{\sqrt{d_k}}$.

\subsubsection{Multi-Head Attention}

Instead of performing a single attention function with $d_{\text{model}}$-dimensional
keys, values and queries, we found it beneficial to linearly project the queries,
keys and values $h$ times with different, learned linear projections to $d_k$,
$d_k$ and $d_v$ dimensions, respectively. On each of these projected versions of
queries, keys and values we then perform the attention function in parallel,
yielding $d_v$-dimensional output values. These are concatenated and once again
projected, resulting in the final values, as depicted in Figure~2.

Multi-head attention allows the model to jointly attend to information from
different representation subspaces at different positions. With a single
attention head, averaging inhibits this.
\[
  \mathrm{MultiHead}(Q, K, V) = \mathrm{Concat}(\mathrm{head}_1, \ldots, \mathrm{head}_h)\,W^O
\]
where $\mathrm{head}_i = \mathrm{Attention}(QW_i^Q,\, KW_i^K,\, VW_i^V)$

Where the projections are parameter matrices
$W_i^Q \in \mathbb{R}^{d_{\text{model}} \times d_k}$,
$W_i^K \in \mathbb{R}^{d_{\text{model}} \times d_k}$,
$W_i^V \in \mathbb{R}^{d_{\text{model}} \times d_v}$ and
$W^O \in \mathbb{R}^{hd_v \times d_{\text{model}}}$.

In this work we employ $h = 8$ parallel attention layers, or heads. For each
of these we use $d_k = d_v = d_{\text{model}}/h = 64$. Due to the reduced
dimension of each head, the total computational cost is similar to that of
single-head attention with full dimensionality.

\subsubsection{Applications of Attention in our Model}

The Transformer uses multi-head attention in three different ways:

\begin{itemize}
  \item In ``encoder-decoder attention'' layers, the queries come from the
    previous decoder layer, and the memory keys and values come from the output
    of the encoder. This allows every position in the decoder to attend over all
    positions in the input sequence. This mimics the typical encoder-decoder
    attention mechanisms in sequence-to-sequence models such as
    \cite{wu2016google,bahdanau2014neural,decomposableattentionmodel}.
  \item The encoder contains self-attention layers. In a self-attention layer
    all of the keys, values and queries come from the same place, in this case,
    the output of the previous layer in the encoder. Each position in the
    encoder can attend to all positions in the previous layer of the encoder.
  \item Similarly, self-attention layers in the decoder allow each position in
    the decoder to attend to all positions in the decoder up to and including
    that position. We need to prevent leftward information flow in the decoder
    to preserve the auto-regressive property. We implement this inside of
    scaled dot-product attention by masking out (setting to $-\infty$) all
    values in the input to the softmax which correspond to illegal connections.
\end{itemize}

\subsection{Position-wise Feed-Forward Networks}

In addition to attention sub-layers, each of the layers in our encoder and
decoder contains a fully connected feed-forward network, which is applied to
each position separately and identically. This consists of two linear
transformations with a ReLU activation in between.
\[
  \mathrm{FFN}(x) = \max(0,\, xW_1 + b_1)\,W_2 + b_2
\]
While the linear transformations are the same across different positions, they
use different parameters from layer to layer. Another way of describing this is
as two convolutions with kernel size 1. The dimensionality of input and output
is $d_{\text{model}} = 512$, and the inner-layer has dimensionality
$d_{ff} = 2048$.

\subsection{Embeddings and Softmax}

Similarly to other sequence transduction models, we use learned embeddings to
convert the input tokens and output tokens to vectors of dimension
$d_{\text{model}}$. We also use the usual learned linear transformation and
softmax function to convert the decoder output to predicted next-token
probabilities. In our model, we share the same weight matrix between the two
embedding layers and the pre-softmax linear transformation, similar to
\cite{press2016using}. In the embedding layers, we multiply those weights by
$\sqrt{d_{\text{model}}}$.

\subsection{Positional Encoding}

Since our model contains no recurrence and no convolution, in order for the
model to make use of the order of the sequence, we must inject some information
about the relative or absolute position of the tokens in the sequence. To this
end, we add ``positional encodings'' to the input embeddings at the bottoms of
the encoder and decoder stacks. The positional encodings have the same dimension
$d_{\text{model}}$ as the embeddings, so that the two can be summed. There are
many choices of positional encodings, learned and fixed \cite{gehring2017convolutional}.

In this work, we use sine and cosine functions of different frequencies:
\[
  PE_{(pos,\,2i)}   = \sin\!\left(\frac{pos}{10000^{2i/d_{\text{model}}}}\right)
  \qquad
  PE_{(pos,\,2i+1)} = \cos\!\left(\frac{pos}{10000^{2i/d_{\text{model}}}}\right)
\]
where $pos$ is the position and $i$ is the dimension. That is, each dimension
of the positional encoding corresponds to a sinusoid. The wavelengths form a
geometric progression from $2\pi$ to $10000 \cdot 2\pi$. We chose this function
because we hypothesized it would allow the model to easily learn to attend by
relative positions, since for any fixed offset $k$,
$PE_{pos+k}$ can be represented as a linear function of $PE_{pos}$.

We also experimented with using learned positional embeddings
\cite{gehring2017convolutional} instead, and found that the two versions
produced nearly identical results (see Table~3 row (E)). We chose the sinusoidal
version because it may allow the model to extrapolate to sequence lengths longer
than the ones encountered during training.

\section{Why Self-Attention}

In this section we compare various aspects of self-attention layers to the
recurrent and convolutional layers commonly used for mapping one variable-length
sequence of symbol representations $(x_1, \ldots, x_n)$ to another sequence of
equal length $(z_1, \ldots, z_n)$, with $x_i, z_i \in \mathbb{R}^d$, such as a
hidden layer in a typical sequence transduction encoder or decoder. Motivating
our use of self-attention we consider three desiderata.

One is the total computational complexity per layer. Another is the amount of
computation that can be parallelized, as measured by the minimum number of
sequential operations required.

The third is the path length between long-range dependencies in the network.
Learning long-range dependencies is a key challenge in many sequence
transduction tasks. One key factor affecting the ability to learn such
dependencies is the length of the paths forward and backward signals have to
traverse in the network. The shorter these paths between any combination of
positions in the input and output sequences, the easier it is to learn
long-range dependencies \cite{hochreiter2001gradient}. Hence we also compare
the maximum path length between any two input and output positions in networks
composed of the different layer types.

As noted in Table~1, a self-attention layer connects all positions with a
constant number of sequentially executed operations, whereas a recurrent layer
requires $O(n)$ sequential operations. In terms of computational complexity,
self-attention layers are faster than recurrent layers when the sequence length
$n$ is smaller than the representation dimensionality $d$, which is most often
the case with sentence representations used by state-of-the-art models in
machine translations, such as word-piece \cite{wu2016google} and byte-pair
\cite{sennrich2015neural} representations.

To improve computational performance for tasks involving very long sequences,
self-attention could be restricted to considering only a neighborhood of size
$r$ in the input sequence centered around the respective output position. This
would increase the maximum path length to $O(n/r)$. We plan to investigate this
approach further in future work.

A single convolutional layer with kernel width $k < n$ does not connect all
pairs of input and output positions. Doing so requires a stack of $O(n/k)$
convolutional layers in the case of contiguous kernels, or $O(\log_k(n))$ in the
case of dilated convolutions \cite{yu2015multi}, increasing the length of the
longest paths between any two positions in the network. Convolutional layers are
generally more expensive than recurrent layers, by a factor of $k$. Separable
convolutions \cite{chollet2016xception}, however, decrease the complexity
considerably, to $O(k \cdot n \cdot d + n \cdot d^2)$. Even with $k = n$,
however, the complexity of a separable convolution is equal to the combination
of a self-attention layer and a point-wise feed-forward layer, the approach we
take in our model.

As side benefit, self-attention could yield more interpretable models. We
inspect attention distributions from our models and present and discuss examples
in the appendix. Not only do individual attention heads clearly learn to perform
different tasks, many appear to exhibit behavior related to the syntactic and
semantic structure of the sentences.

\section{Training}

This section describes the training regime for our models.

\subsection{Training Data and Batching}

We trained on the standard WMT 2014 English-German dataset consisting of about
4.5 million sentence pairs. Sentences were encoded using byte-pair encoding
\cite{sennrich2015neural}, which has a shared source-target vocabulary of about
37000 tokens. For English-French, we used the significantly larger WMT 2014
English-French dataset consisting of 36M sentences and split tokens into a
32000 word-piece vocabulary \cite{wu2016google}. Sentence pairs were batched
together by approximate sequence length. Each training batch contained a set of
sentence pairs containing approximately 25000 source tokens and 25000 target
tokens.

\subsection{Hardware and Schedule}

We trained our models on one machine with 8 NVIDIA P100 GPUs. For our base
models using the hyperparameters described throughout the paper, each training
step took about 0.4 seconds. We trained the base models for a total of 100,000
steps or 12 hours. For our big models, step time was 1.0 seconds. The big
models were trained for 300,000 steps (3.5 days).

\subsection{Optimizer}

We used the Adam optimizer \cite{kingma2014adam} with $\beta_1 = 0.9$,
$\beta_2 = 0.98$ and $\epsilon = 10^{-9}$. We varied the learning rate over the
course of training, according to the formula:
\[
  lrate = d_{\text{model}}^{-0.5} \cdot
  \min\!\left(step\_num^{-0.5},\;
  step\_num \cdot warmup\_steps^{-1.5}\right)
\]
This corresponds to increasing the learning rate linearly for the first
$warmup\_steps$ training steps, and decreasing it thereafter proportionally to
the inverse square root of the step number. We used $warmup\_steps = 4000$.

\subsection{Regularization}

We employ three types of regularization during training:

\textbf{Residual Dropout.} We apply dropout \cite{srivastava2014dropout} to the
output of each sub-layer, before it is added to the sub-layer input and
normalized. In addition, we apply dropout to the sums of the embeddings and the
positional encodings in both the encoder and decoder stacks. For the base model,
we use a rate of $P_{drop} = 0.1$.

\textbf{Label Smoothing.} During training, we employed label smoothing of value
$\epsilon_{ls} = 0.1$ \cite{szegedy2015rethinking}. This hurts perplexity, as
the model learns to be more unsure, but improves accuracy and BLEU score.

\section{Results}

\subsection{Machine Translation}

On the WMT 2014 English-to-German translation task, the big transformer model
(Transformer (big) in Table~2) outperforms the best previously reported models
including ensembles by more than 2.0 BLEU, establishing a new
state-of-the-art BLEU score of 28.4. The configuration of this model is listed
in the bottom line of Table~3. Training took 3.5 days on 8 P100 GPUs. Even our
base model surpasses all previously published models and ensembles, at a
fraction of the training cost of any of the competitive models.

On the WMT 2014 English-to-French translation task, our big model achieves a
BLEU score of 41.0, outperforming all of the previously published single models,
at less than $1/4$ the training cost of the previous state-of-the-art model.
The Transformer (big) model trained for English-to-French used dropout rate
$P_{drop} = 0.1$, instead of 0.3.

For the base models, we used a single model obtained by averaging the last 5
checkpoints, which were written at 10-minute intervals. For the big models, we
averaged the last 20 checkpoints. We used beam search with a beam size of 4 and
length penalty $\alpha = 0.6$ \cite{wu2016google}. These hyperparameters were
chosen after experimentation on the development set. We set the maximum output
length during inference to input length $+ 50$, but terminate early when
possible \cite{wu2016google}.

Table~2 summarizes our results and compares our translation quality and training
costs to other model architectures from the literature. We estimate the number
of floating point operations used to train a model by multiplying the training
time, the number of GPUs used, and an estimate of the sustained single-precision
floating-point capacity of each GPU.

\subsection{Model Variations}

To evaluate the importance of different components of the Transformer, we varied
our base model in different ways, measuring the change in performance on
English-to-German translation on the development set, newstest2013. We used beam
search as described in the previous section, but no checkpoint averaging. We
present these results in Table~3.

In Table~3 rows (A), we vary the number of attention heads and the attention key
and value dimensions, keeping the amount of computation constant, as described
in Section~3.2.2. While single-head attention is 0.9 BLEU worse than the best
setting, quality also drops off with too many heads.

In Table~3 rows (B), we observe that reducing the attention key size $d_k$ hurts
model quality. This suggests that determining compatibility is not easy and that
a more sophisticated compatibility function than dot product may be beneficial.
We further observe in rows (C) and (D) that, as expected, bigger models are
better, and dropout is very helpful in avoiding over-fitting. In row (E) we
replace our sinusoidal positional encoding with learned positional embeddings
\cite{gehring2017convolutional}, and observe nearly identical results to the
base model.

\subsection{English Constituency Parsing}

To evaluate if the Transformer can generalize to other tasks we performed
experiments on English constituency parsing. This task presents specific
challenges: the output is subject to strong structural constraints and is
significantly longer than the input. Furthermore, RNN sequence-to-sequence
models have not been able to attain state-of-the-art results in small-data
regimes \cite{vinyals2015grammar}.

We trained a 4-layer transformer with $d_{\text{model}} = 1024$ on the Wall
Street Journal (WSJ) portion of the Penn Treebank \cite{marcus1993building},
about 40K training sentences. We also trained it in a semi-supervised setting,
using the larger high-confidence and BerkleyParser corpora from with approximately
17M sentences \cite{vinyals2015grammar,zhu2013fast}. We used a vocabulary of
16K tokens for the WSJ only setting and a vocabulary of 32K tokens for the
semi-supervised setting.

We performed only a small number of experiments to select the dropout, both
attention and residual (section~5.4), learning rates and beam size on the
Section 22 development set, all other parameters remained unchanged from the
English-to-German base translation model. During inference, we increased the
maximum output length to input length $+ 300$. We used a beam size of 21 and
$\alpha = 0.3$ for both WSJ only and the semi-supervised setting.

Our results in Table~4 show that despite the lack of task-specific tuning our
model performs surprisingly well, yielding better results than all previously
reported models with the exception of the Recurrent Neural Network Grammar
\cite{dyer2016recurrent}.

In contrast to RNN sequence-to-sequence models \cite{vinyals2015grammar}, the
Transformer outperforms the BerkeleyParser \cite{petrov2006learning} even when
training only on the WSJ training set of 40K sentences.

\section{Conclusion}

In this work, we presented the Transformer, the first sequence transduction model
based entirely on attention, replacing the recurrent layers most commonly used in
encoder-decoder architectures with multi-headed self-attention.

For translation tasks, the Transformer can be trained significantly faster than
architectures based on recurrent or convolutional layers. We achieved a new state
of the art on both WMT 2014 English-to-German and WMT 2014 English-to-French
translation tasks. In the former task our best model outperforms even all
previously reported ensembles.

We are excited about the future of attention-based models and plan to apply them
to other tasks. We plan to extend the Transformer to problems involving input and
output modalities other than text and to investigate local, restricted attention
mechanisms to efficiently handle large inputs and outputs such as images, audio
and video. Making generation less sequential is another research goal of ours.

The code we used to train and evaluate our models is available at
\url{https://github.com/tensorflow/tensor2tensor}.
